Large and Infinitary Quotient Inductive-Inductive Types

### 2019-12-30 draft/paper.tex
paper

\input{abbrevs.tex}
\newcommand{\Lift}{\Uparrow}
\newcommand{\ToS}{\mathsf{ToS}}
\newcommand{\ext}{\triangleright}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\Pii}{\Pi}
\newcommand{\appi}{\mathsf{app}}
\newcommand{\lami}{\mathsf{lam}}
\newcommand{\Pie}{\hat{\Pi}}
\newcommand{\appe}{\mathsf{ap\hat{p}}}
\newcommand{\lame}{\mathsf{la\hat{m}}}
\newcommand{\Piinf}{\tilde{\Pi}}
\newcommand{\appinf}{\mathsf{ap\tilde{p}}}
\newcommand{\laminf}{\mathsf{la\tilde{m}}}
\newcommand{\appitt}{\mathop{{\scriptstyle @}}}
\newcommand{\Refl}{\mathsf{Refl}}
\newcommand{\IdU}{\mathsf{IdU}}
\newcommand{\ReflU}{\mathsf{ReflU}}
\newcommand{\Sig}{\mathsf{Sig}}
\newcommand{\Sg}{\Sigma}
\newcommand{\flCwF}{\mathsf{flCwF}}
\newcommand{\Kfam}{\mathsf{K}}
\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}
\newcommand{\lamK}{\mathsf{lam}^{\K}}
\newcommand{\appK}{\mathsf{app}^{\K}}

\newcommand{\bCon}{\boldsymbol{\Con}}
\newcommand{\bTy}{\boldsymbol{\Ty}}
\newcommand{\bSub}{\boldsymbol{\Sub}}
\newcommand{\bTm}{\boldsymbol{\Tm}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bsigma}{\boldsymbol{\sigma}}
\newcommand{\bdelta}{\boldsymbol{\delta}}
\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
\newcommand{\bt}{\boldsymbol{t}}
\newcommand{\bu}{\boldsymbol{u}}
\newcommand{\bA}{\boldsymbol{A}}
\newcommand{\bB}{\boldsymbol{B}}
\newcommand{\bid}{\boldsymbol{\id}}
\newcommand{\bemptycon}{\boldsymbol{\emptycon}}

@@

Signatures are given as contexts in a certain type theory, the theory of
signatures. We shall abbreviate it as ToS. However, ToS turns out to be a large
infinitary QIIT itself, and we would like to define ToS and a notion of
signature without referring to QIITs, only using features present in cETT. As a
first step, we define a notion of model. $\Pi$

@@

A \textbf{category with families} (CwF), where all four underlying sets (of
objects, morphisms, types and terms) are in $\Set_j$. Following notation in
\cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$,
$\Sub : \Con \ra \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma :
\Con) \ra \Ty\,\Gamma \ra \Set_j$. We use $\id$ and $\blank\circ\blank$ to
denote identity and composition for substitution. We denote the empty context as
$\emptycon : \Con$, and the unique substitution into the empty context as
$\epsilon : \Sub\,\Gamma\,\emptycon$. Context extension is $\blank\ext\blank :
(\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$. Substitution on types and terms is
written as $\blank[\blank]$. Projections are noted as $\p :
\Sub\,(\Gamma\,\ext\,A)\,\Gamma$ and $\q : \Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and
substitution extension is $\blank,\blank : (\sigma : \Sub\,\Gamma\,\Delta)\ra
\Tm\,\Gamma\,A[\sigma]\ra \Sub\,\Gamma\,(\Delta\ext A)$.

@@

\textbf{Inductive function space}, with $\Pii : (a : \Tm\,\Gamma\,\U) \ra
\Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$ and application as $\appi :
\Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$ and its
inverse $\lami$.

\textbf{External function space} with $\Set_i$ domain: $\Pie : (A : \Set_i)\ra(A
\ra \Ty\,\Gamma)\ra \Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x
: A)\ra \Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.

\textbf{Infinitary function space}, with $\Piinf : (A : \Set_i)\ra(A \ra
\Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, $\appinf :
\Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
$\laminf$.

\textbf{An identity type} $\Id : (a : \Tm\,\Gamma\,\U)\ra
\Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
(T : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.

@@

\emph{Notational conventions.} We name elements of $\Con$ as $\Gamma$, $\Delta$,
$\Theta$, elements of $\Sub\,\Gamma\,\Delta$ as $\sigma$, $\delta$, $\nu$,
elements of $\Ty\,\Gamma$ as $A$, $B$, $C$, and elements of $\Tm\,\Gamma\,A$ as
$t$, $u$, $v$. CwF components by default support de Bruijn indices, which are
not easily readable. We use instead a nameful notation for binders in context
extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
\El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
\begin{alignat*}{3}
  & \blank\appitt\blank && :
      \Tm\,\Gamma\,(\Pii\,a\,B)\ra
      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
  & t\appitt u && :\equiv (\appi\,t)[\id, u]
\end{alignat*}

@@

\begin{example}{
    Signature for natural numbers. Here, no external types appear, so the level
    can be chosen as $0$.}
\begin{alignat*}{3}
  & NatSig : \Sig_0 \\
  & NatSig :\equiv \lambda (j :\mathbb{N})(M : \ToS_{0,j}).\\
  & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
  & \hspace{2.67em}\ext_M (suc : \Pii_M\,N\,(\El_M\,N)))
\end{alignat*}
\end{example}

@@

\begin{example}{Infinitary constructors}. The universe $\U$ is closed under
the $\Piinf$ function type, which allows such functions to appear in the domains
of $\Pii$ types. This allows a signature for trees branching with arbitrary
small sets. This is a signature at level 1, since we have $\Set_0$ as a $\Pie$
domain type.
\[
TreeSig :\equiv
\emptycon \ext (Tree : \U) \ext (node : \Pie\,\Set_0\,\lambda A.\,\Pii\,(\Piinf\,A\,\lambda \_.\,Tree)\,(\El\,Tree))
\]
\end{example}

@@

\begin{example}{Recursive equations}. Again, the universe is closed under $\Id$, which allows us to write equations
in $\Pii$ domains. A minimal example:
\[
RecEqSig :\equiv \emptycon \ext (A : \U) \ext (a : \El\,A) \ext (f : \Pii\,(x : A)\,(\Pii\,(\Id\,A\,x\,a)\,(\El\,A)))
\]
More interesting (and complicated) examples are boundary conditions in various
cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated equations
as well, but these are all trivial in the semantics, where we assume UIP.
\end{example}

@@

We make a similar choice in the interpretation of signatures themselves: we
use structured CwFs instead of lex categories to model notions of induction. The
reason here is that CwFs allow us to define induction in strictly the same way as
one would write it down in type theory, since we have $\Ty$ and $\Tm$ for a
primitive notion of dependent objects and morphisms. In contrast, dependent
objects in lex categories is a derived notion, and the induction principles we
get are only up to isomorphisms. This issue is perhaps not relevant from a
purely categorical perspective, but we are concerned with eventually
implementing QIITs in proof assistants, so we prefer if our semantics computes
strictly.

@@

@@

We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$. We define
$\flCwF$ as follows.

\begin{definition}[Finite limit CwFs]\label{def:flCwF}
For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
iterated $\Sigma$-type with the following components:
\begin{enumerate}
  \item A CwF with underlying sets all in $\Set_i$. We reuse the notation
    from definition \ref{def:tos}.
  \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
    \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
  \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with $\refl$, equality reflection and UIP.
  \item Constant families. This includes a type former $\Kfam : \Con \ra
    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, together with $\lamK
    : \Sub\,\Gamma\,\Delta \ra \Tm\,\Gamma\,(\Kfam\,\Delta)$ and its inverse
    $\appK$. The idea is that $\Kfam\,\Delta$ is a representation of $\Delta$ as
    a type in any context. Clairambault and Dybjer called constant families
    ``democracy'' in \cite{clairambault2014biequivalence}.
\end{enumerate}
\end{definition}

@@

We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.

@@

\subsection{Type Substitution}
We aim to define $\boldsymbol{\blank[\blank] : \Ty\,\Delta \ra
  \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$, such that $\boldsymbol{A[\id]} = \bA$
and $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$. This
is rather laborious to fully write out, so we omit most details and focus on the
main points.

The underlying sets are given by simple composition:
\begin{alignat*}{3}
  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && :\equiv \Con_{\bA}\,(\bsigma\,\ulGamma)\\
  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :\equiv
    \Sub_{\bA}\,\Gamma\,\Delta\,(\bsigma\,\ulsigma)\\
  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :\equiv
      \Ty_{\bA}\,\Gamma\,(\bsigma\,\ulA)\\
  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :\equiv
      \Tm_{\bA}\,\Gamma\,A\,(\bsigma\,\ult)
\end{alignat*}

@@

Functoriality of type substitution, i.e.\ $\boldsymbol{A[\id]} = \bA$ and
$\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follow
from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe_{\id}$,
$\coe_{\circ}$, $\coh_{\id}$ and $\coh_{\circ}$ in $\bA$.

@@

$\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of \emph{weak flCwF sections} of $\bA$.

\begin{definition}[Weak flCwF section] Given $\bA : \bTy\,\bGamma$, the underlying functions of $\bt : \bTm\,\bGamma\,\bA$ are as follows:
\begin{alignat*}{3}
  & \bt : (\ulGamma : \Con_{\bGamma}) \ra \Con_{\bA}\,\ulGamma\\
  & \bt : (\ulsigma : \Sub_{\bGamma}\,\ulGamma\,\ulDelta)
         \ra \Sub_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulDelta)\,\ulsigma\\
  & \bt : (\ulA : \Ty_{\bGamma}) \ra \Ty_{\bA}\,(\bt\,\ulGamma)\,\ulA\\
  & \bt : (\ult : \Tm_{\bGamma}\,\ulGamma\,\ulA) \ra
          \Tm_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulA)\,\ult
\end{alignat*}
Such that
\begin{enumerate}
  \item $\bsigma\,(\ulA[\ulsigma]) = (\bsigma\,\ulA)\,[\bsigma\,\ulsigma]$
  \item $\bsigma\,(\ult[\ulsigma]) = (\bsigma\,\ult)\,[\bsigma\,\ulsigma]$
  \item The unique map $\epsilon_{\bA} : \Sub\,(\bsigma\,\emptycon)\,\emptycon_{\bGamma}\,\id$ has a vertical retraction.
  \item Each $(\bsigma\,\p,\,\bsigma\,\q) : \Sub\,(\bsigma\,(\ulGamma\,\ext\,\ulA))\,(\bsigma\,\ulGamma\,\ext\,\bsigma\,\ulA)\,\id$ has a vertical inverse.
\end{enumerate}
Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
isomorphisms as $\bt_{\emptycon} : \bt\,\emptycon \simeq_{\id} \emptycon$ and
$\bt_{\ext} : \bt\,(\ulGamma\ext \ulA) \simeq_{\id} \bt\,\ulGamma \ext
\bt\,\ulA$. In short, weak section is a dependently typed analogue of weak
morphism, with dependent underlying functions and displayed preservation
isomorphisms. We also have the derived fl-preservation isomorphisms.
\end{definition}

@@

$\boldsymbol{\blank[\blank] : \Tm\,\Delta\,A \ra (\sigma : \Sub\,\Gamma\,\Delta)
  \ra \Tm\,\Gamma\,(A[\sigma])}$ is given the similarly to
$\boldsymbol{\blank\circ\blank}$ in Section \ref{TODO}. Underlying functions
are given by function composition, and preservation morphisms are also similar:
\begin{alignat*}{3}
  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :\equiv
    \bt\,\bsigma_{\emptycon}^{-1} \circ \bt_{\emptycon}^{-1} \\
  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :\equiv
    \bt\,\bsigma_{\ext}^{-1} \circ \bt_{\ext}^{-1}
\end{alignat*}

@@

$\boldsymbol{\Gamma \ext A : \Con}$ is defined as the \emph{total flCwF} of
$\bA$. This is given by bundling together all displayed flCwF components in
$\bA$ with corresponding base components in $\bGamma$, using metatheoretic
$\Sigma$-types. It is a straightforward extension of total categories in
\cite{displayedcats}.

$\boldsymbol{\p : \Sub\,(\Gamma\ext A)\,\Gamma}$ is a strict morphism given by
taking a first projection for each component. $\boldsymbol{\q : \Tm\,(\Gamma\ext
  A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections.

>>>>

\input{abbrevs.tex}
\newcommand{\Lift}{\Uparrow}
\newcommand{\ToS}{\mathsf{ToS}}
\newcommand{\ext}{\triangleright}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\Pii}{\Pi}
\newcommand{\appi}{\mathsf{app}}
\newcommand{\lami}{\mathsf{lam}}
\newcommand{\Pie}{\Pi^{\mathsf{ext}}}
\newcommand{\appe}{\mathsf{app^{ext}}}
\newcommand{\lame}{\mathsf{lam^{ext}}}
\newcommand{\Piinf}{\Pi^{\mathsf{inf}}}
\newcommand{\appinf}{\mathsf{app^{inf}}}
\newcommand{\laminf}{\mathsf{lam^{inf}}}
\newcommand{\appitt}{\mathop{{\scriptstyle @}}}
\newcommand{\Refl}{\mathsf{Refl}}
\newcommand{\IdU}{\mathsf{IdU}}
\newcommand{\ReflU}{\mathsf{ReflU}}
\newcommand{\Sig}{\mathsf{Sig}}
\newcommand{\Sg}{\Sigma}
\newcommand{\flCwF}{\mathsf{flCwF}}
\newcommand{\Kfam}{\mathsf{K}}
\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}
\newcommand{\lamK}{\mathsf{lam}^{\K}}
\newcommand{\appK}{\mathsf{app}^{\K}}

\newcommand{\arri}{\Rightarrow}
\newcommand{\arre}{\Rightarrow^{\mathsf{ext}}}
\newcommand{\arrinf}{\Rightarrow^{\mathsf{inf}}}

\newcommand{\bCon}{\boldsymbol{\Con}}
\newcommand{\bTy}{\boldsymbol{\Ty}}
\newcommand{\bSub}{\boldsymbol{\Sub}}
\newcommand{\bTm}{\boldsymbol{\Tm}}
\newcommand{\bGamma}{\boldsymbol{\Gamma}}
\newcommand{\bDelta}{\boldsymbol{\Delta}}
\newcommand{\bsigma}{\boldsymbol{\sigma}}
\newcommand{\bdelta}{\boldsymbol{\delta}}
\newcommand{\bepsilon}{\boldsymbol{\epsilon}}
\newcommand{\bt}{\boldsymbol{t}}
\newcommand{\bu}{\boldsymbol{u}}
\newcommand{\bA}{\boldsymbol{A}}
\newcommand{\ba}{\boldsymbol{a}}
\newcommand{\bB}{\boldsymbol{B}}
\newcommand{\bid}{\boldsymbol{\id}}
\newcommand{\bemptycon}{\boldsymbol{\emptycon}}
\newcommand{\bSet}{\boldsymbol{\Set}}
\newcommand{\bU}{\boldsymbol{\U}}
\newcommand{\bEl}{\boldsymbol{\El}}
\newcommand{\bPii}{\boldsymbol{\Pi}}
\newcommand{\bappi}{\boldsymbol{\mathsf{app}}}
\newcommand{\blami}{\boldsymbol{\mathsf{lam}}}

@@

Signatures are given as contexts in a certain type theory, the theory of
signatures. We shall abbreviate it as ToS. However, ToS turns out to be a large
infinitary QIIT itself, and we would like to define ToS and a notion of
signature without referring to QIITs, only using features present in cETT. As a
first step, we define a notion of model.

@@

A \textbf{category with families} (CwF), where all four underlying sets (of
objects, morphisms, types and terms) are in $\Set_j$. Following notation in
\cite{kaposi2019constructing}, we denote these respectively as $\Con : \Set_j$,
$\Sub : \Con \ra \Con \ra \Set_j$, $\Ty : \Con \ra \Set_j$ and $\Tm : (\Gamma :
\Con) \ra \Ty\,\Gamma \ra \Set_j$. We use $\id$ and $\blank\circ\blank$ to
denote identity and composition for substitution. We denote the empty context as
$\emptycon : \Con$, and the unique substitution into the empty context as
$\epsilon : \Sub\,\Gamma\,\emptycon$. Context extension is $\blank\ext\blank :
(\Gamma : \Con)\ra \Ty\,\Gamma \ra \Con$. Substitution on types and terms is
written as $\blank[\blank]$. Projections are noted as $\p :
\Sub\,(\Gamma\,\ext\,A)\,\Gamma$ and $\q : \Tm\,(\Gamma\,\ext\,A)\,(A[\p])$, and
substitution extension is $\blank,\blank : (\sigma : \Sub\,\Gamma\,\Delta)\ra
\Tm\,\Gamma\,(A[\sigma])\ra \Sub\,\Gamma\,(\Delta\ext A)$.

@@

\textbf{Inductive function space} $\Pii : (a : \Tm\,\Gamma\,\U) \ra
\Ty\,(\Gamma\,\ext\,\El\,a) \ra \Ty\,\Gamma$, with application as $\appi :
\Tm\,\Gamma\,(\Pii\,a\,B)\ra \Tm\,(\Gamma\,\ext\,\El\,a)\,B$ and its
inverse $\lami$.

\textbf{External function space} $\Pie : (A : \Set_i)\ra(A \ra \Ty\,\Gamma)\ra
\Ty\,\Gamma$, with $\appe : \Tm\,\Gamma\,(\Pie\,A\,B)\ra((x : A)\ra
\Tm\,\Gamma\,(B\,x))$ and its inverse $\lame$.

\textbf{Infinitary function space} $\Piinf : (A : \Set_i)\ra(A \ra
\Tm\,\Gamma\,\U)\ra \Tm\,\Gamma\,\U$, with $\appinf :
\Tm\,\Gamma\,(\El\,(\Pie\,A\,b))\ra((x : A)\ra \Tm\,\Gamma\,(\El\,(b\,x)))$ and
its inverse $\laminf$.

\textbf{An identity type} $\Id : (a : \Tm\,\Gamma\,\U)\ra
\Tm\,\Gamma\,(\El\,a)\ra\Tm\,\Gamma\,(\El\,a)\ra \Tm\,\Gamma\,\U$, with $\Refl :
(t : \Tm\,\Gamma\,(\El\,a))\ra \Id\,a\,t\,t$, equality reflection and UIP.

@@

\emph{Notational conventions.} We name elements of $\Con$ as $\Gamma$, $\Delta$,
$\Theta$, elements of $\Sub\,\Gamma\,\Delta$ as $\sigma$, $\delta$, $\nu$,
elements of $\Ty\,\Gamma$ as $A$, $B$, $C$, and elements of $\Tm\,\Gamma\,A$ as
$t$, $u$, $v$. CwF components by default support de Bruijn indices, which are
not easily readable. We use instead a nameful notation for binders in context
extension, $\Pii$ and $\lami$, e.g.\ as $(\emptycon\,\ext (a : U) \ext (t :
\El\,a))$. We also define a type-theoretic flavor of $\appi$ for convenience:
\begin{alignat*}{3}
  & \blank\appitt\blank && :
      \Tm\,\Gamma\,(\Pii\,a\,B)\ra
      (u : \Tm\,\Gamma\,(\El\,a)) \ra \Tm\,\Gamma\,(B[\id, u])\\
  & t\appitt u && :\equiv (\appi\,t)[\id, u]
\end{alignat*}
We abbreviate non-dependent inductive $\Pii$ as $\blank\arri\blank$, and likewise we use $\blank\arre\blank$ and $\blank\arrinf\blank$.

%% We use abbreviations for non-dependent function types:
%% \begin{alignat*}{3}
%%   & A\arri B && :\equiv \Pi\,A\,(B[\p])\\
%%   & A \arre\,B &&:\equiv \Pie\,A\,(\lambda\,\_.\,B) \\
%%   & A \arrinf\,B &&:\equiv \Piinf\,A\,(\lambda\,\_.\,B)
%% $A\arri B :\equiv \Pi\,A\,(B[\p])$, and $A \arre\,B :\equiv \Pie\,A\,(\lambda\,\_.\,B)$, and likewise
%% $A \arrinf\,B :\equiv \Piinf\,A\,(\lambda\,\_.\,B)$.

@@

\begin{example}{
    Signature for natural numbers. Here, no external types appear, so the level
    can be chosen as $0$.}
\begin{alignat*}{3}
  & NatSig : \Sig_0 \\
  & NatSig :\equiv \lambda (j :\mathbb{N})(M : \ToS_{0,j}).\\
  & \hspace{1em}(\emptycon_M\,\ext_M\, (N : \U_M) \,\ext_M\,(zero : \El_M\,N)\\
  & \hspace{2.67em}\ext_M (suc : N\arri_M\El_M\,N))
\end{alignat*}
\end{example}

@@

In the following examples, we leave the abstracted $M : \ToS_{0,j}$ implicit.

\begin{example}{Infinitary constructors}. The universe $\U$ is closed under
the $\Piinf$ function type, which allows such functions to appear in the domains
of $\Pii$ types. This allows, for example, a signature for trees branching with
arbitrary small sets. This is a signature at level 1, since we have $\Set_0$ as
a $\Pie$ domain type.
\begin{alignat*}{3}
& TreeSig :\equiv \\
& \hspace{1em}\emptycon \ext (Tree : \U)\\
& \hspace{1.35em}\ext (node : \Pie\,\Set_0\,(\lambda A.\,(A\,\arrinf\,Tree)\arri\El\,Tree))
\end{alignat*}
\end{example}

@@

\begin{example}{Recursive equations}. Again, the universe is closed under $\Id$, which allows us to write equations
in $\Pii$ domains. A minimal example:
\begin{alignat*}{3}
  & RecEqSig :\equiv\\
  & \hspace{0.5em}\emptycon\,\ext (A : \U) \ext (a : \El\,A)\ext (f : \Pii\,(x : A)\,(\Id\,A\,x\,a\arri\El\,A))
\end{alignat*}
More interesting (and complicated) examples are boundary conditions in various
cubical type theories\cite{TODO}. Note that our $\Id$ allows iterated equations
as well, but these are all trivial in the semantics, where we assume UIP.
\end{example}

@@

We make a similar choice in the interpretation of signatures themselves: we use
structured CwFs instead of lex categories to model notions of induction. The
reason here is that CwFs allow us to compute induction principles in strictly
the same way as one would write them down in type theory, since we have $\Ty$
and $\Tm$ for a primitive notion of dependent objects and morphisms. In
contrast, dependent objects in lex categories is a derived notion, and the
induction principles we get are only up to isomorphism. This issue is perhaps
not relevant from a purely categorical perspective, but we are concerned with
eventually implementing QIITs in proof assistants, so we prefer if our semantics
computes strictly.

@@

The model involves a large amount of technical detail, most of which we omit,
and only present the most salient parts.

@@

We define $\bCon : \Set_{1+\max(i,j+1)}$ as $\flCwF_{\max(i,j+1)}$.
%% We define
%% $\flCwF$ as follows.

\begin{definition}[Finite limit CwFs]\label{def:flCwF}
For a level $i$, there is a cETT type $\flCwF_i : \Set_{i+1}$, which is an
iterated $\Sigma$-type with the following components:
\begin{enumerate}
  \item A CwF with underlying sets all in $\Set_i$. We reuse the component
    notations from Definition \ref{def:tos}.
  \item $\Sigma$-type $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
    \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
  \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with $\refl$, equality reflection and UIP.
  \item Constant families. This includes a type former $\Kfam : \Con \ra
    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, together with $\lamK
    : \Sub\,\Gamma\,\Delta \ra \Tm\,\Gamma\,(\Kfam\,\Delta)$ and its inverse
    $\appK$. The idea is that $\Kfam\,\Delta$ is a representation of $\Delta$ as
    a type in any context. Clairambault and Dybjer called constant families
    ``democracy'' in \cite{clairambault2014biequivalence}.
\end{enumerate}
\end{definition}

@@

%% We define $\boldsymbol{\Sub\,\Gamma\,\Delta} : \Set_{1 + \max(i, j+1)}$ as the type
%% of weak flCwF morphisms from $\boldsymbol{\Gamma}$ to $\boldsymbol{\Delta}$.

@@

\subsection{Type Substitution}
We aim to define $\boldsymbol{\blank[\blank] : \Ty\,\Delta \ra
  \Sub\,\Gamma\,\Delta \ra \Ty\,\Gamma}$, such that $\boldsymbol{A[\id]} = \bA$
and $\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$. The
underlying sets are given by simple composition:
\begin{alignat*}{3}
  & \Con_{\boldsymbol{A[\sigma]}}\,\ulGamma && :\equiv \Con_{\bA}\,(\bsigma\,\ulGamma)\\
  & \Sub_{\boldsymbol{A[\sigma]}}\,\Gamma\,\Delta\,\ulsigma && :\equiv
    \Sub_{\bA}\,\Gamma\,\Delta\,(\bsigma\,\ulsigma)\\
  & \Ty_{\boldsymbol{A[\sigma]}}\,\Gamma\,\ulA && :\equiv
      \Ty_{\bA}\,\Gamma\,(\bsigma\,\ulA)\\
  & \Tm_{\boldsymbol{A[\sigma]}}\,\Gamma\,A\,\ult && :\equiv
      \Tm_{\bA}\,\Gamma\,A\,(\bsigma\,\ult)
\end{alignat*}

@@

Functoriality of type substitution, i.e.\ $\boldsymbol{A[\id]} = \bA$ and
$\boldsymbol{A[\sigma\circ\delta]} = \boldsymbol{A[\sigma][\delta]}$, follows
from Lemma \ref{lem:idcomppres} and split cleaving given by $\coe_{\id}$,
$\coe_{\circ}$, $\coh_{\id}$ and $\coh_{\circ}$ in $\bA$.

@@

\begin{definition} $\boldsymbol{\Tm\,\Gamma\,A} : \Set_{1 + \max(i, j)}$ is defined as the type of
\emph{weak flCwF sections} of $\bA$. The underlying functions of $\bt :
\bTm\,\bGamma\,\bA$ are as follows:
\begin{alignat*}{3}
  & \bt : (\ulGamma : \Con_{\bGamma}) \ra \Con_{\bA}\,\ulGamma\\
  & \bt : (\ulsigma : \Sub_{\bGamma}\,\ulGamma\,\ulDelta)
         \ra \Sub_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulDelta)\,\ulsigma\\
  & \bt : (\ulA : \Ty_{\bGamma}) \ra \Ty_{\bA}\,(\bt\,\ulGamma)\,\ulA\\
  & \bt : (\ult : \Tm_{\bGamma}\,\ulGamma\,\ulA) \ra
          \Tm_{\bA}\,(\bt\,\ulGamma)\,(\bt\,\ulA)\,\ult
\end{alignat*}
Such that
\begin{enumerate}
  \item $\bt\,(\ulA[\ulsigma]) = (\bt\,\ulA)\,[\bt\,\ulsigma]$
  \item $\bt\,(\ult[\ulsigma]) = (\bt\,\ult)\,[\bt\,\ulsigma]$
  \item The unique map $\epsilon_{\bA} : \Sub\,(\bt\,\emptycon)\,\emptycon\,\id$ has a vertical retraction.
  \item Each $(\bt\,\p,\,\bt\,\q) : \Sub\,(\bt\,(\ulGamma\,\ext\,\ulA))\,(\bt\,\ulGamma\,\ext\,\bt\,\ulA)\,\id$ has a vertical inverse.
\end{enumerate}
\end{definition}
Similarly to Definition \ref{def:weakmorphism}, we denote the evident preservation
isomorphisms as $\bt_{\emptycon} : \bt\,\emptycon \simeq_{\id} \emptycon$ and
$\bt_{\ext} : \bt\,(\ulGamma\ext \ulA) \simeq_{\id} \bt\,\ulGamma \ext
\bt\,\ulA$. In short, weak section is a dependently typed analogue of weak
morphism, with dependent underlying functions and displayed preservation
isomorphisms. We also have the derived fl-preservation isomorphisms.

@@

$\boldsymbol{\blank[\blank] : \Tm\,\Delta\,A \ra (\sigma : \Sub\,\Gamma\,\Delta)
  \ra \Tm\,\Gamma\,(A[\sigma])}$ is given similarly to
$\boldsymbol{\blank\circ\blank}$ in Section \ref{TODO}. Underlying functions
are given by function composition, and preservation morphisms are also similar:
\begin{alignat*}{3}
  & (\boldsymbol{t[\sigma]})_{\emptycon}^{-1} :\equiv
    \bt\,\bsigma_{\emptycon}^{-1} \circ \bt_{\emptycon}^{-1} \\
  & (\boldsymbol{t[\sigma]})_{\ext}^{-1} :\equiv
    \bt\,\bsigma_{\ext}^{-1} \circ \bt_{\ext}^{-1}
\end{alignat*}

@@

$\boldsymbol{\Gamma \ext A : \Con}$ is defined as the \emph{total flCwF} of
$\bA$. This is given by bundling together all displayed flCwF components in
$\bA$ with corresponding base components in $\bGamma$, using the metatheoretic
$\Sigma$-type. It is a straightforward extension of total categories in
\cite{displayedcats}.

$\boldsymbol{\p : \Sub\,(\Gamma\ext A)\,\Gamma}$ is a strict morphism given by
taking a first projection for each component. $\boldsymbol{\q : \Tm\,(\Gamma\ext
  A)\,(A[\p])}$ is likewise a strict flCwF section given by second projections. Substitution extension $\boldsymbol{\sigma,\,t}$ is given by pointwise combining $\bsigma$ and $\bt$ with metatheoretic $\Sigma$ pairing, e.g.\ $\Con_{\boldsymbol{(\sigma,t)}}\,\ulGamma :\equiv (\bsigma\,\ulGamma,\,\bt\,\ulGamma)$.

\subsection{Universe}

\begin{definition} For a level $i$, we write $\bSet_i$ for the flCwF of sets where $\Con_{\bSet_i} :\equiv \Set_i$ and $\Sub_{\bSet_i}\,\Gamma\,\Delta :\equiv \Gamma \ra \Delta$.
\end{definition}

We define $\boldsymbol{\U : \Ty\,\Gamma}$ as the isofibration which is
constantly $\bSet_j$. A constant isofibration does not actually depend on
the base flCwF, and has trivial iso-cleaving where $\coe$-s are identity
functions. Hence, we have $\Con_{\bU}\,\ulGamma :\equiv \Set_j$ and $\Sub_{\bU}\,\Gamma\,\Delta\,\ulsigma :\equiv \Gamma \ra \Delta$.

\emph{Remark.} The type $\bTm\,\bGamma\,\bU$ is strictly equal to $\bSub\,\bGamma\,\bSet_j$, so in the following we will think about semantic elements of the universe as weak morphisms from $\bGamma$ to $\bSet_j$.

\subsection{Elements of the Universe}
